PyTracer: Automatically Profiling Numerical Instabilities in Python

### 2022-04-14 IEEEtran/paper.tex
Revision

\usepackage{todonotes}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{color}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{xcolor,colortbl}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue!50!black,citecolor=blue!75!black,filecolor=black,urlcolor=blue!75!black}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage{makecell}
\usepackage{ulem} % provides sout
\usepackage[frozencache=true,cachedir=minted-cache]{minted} 
\usepackage[htt]{hyphenat}      
\usepackage{subcaption}
\usepackage{graphicx,import}
\usepackage{booktabs}
\usepackage{soul}

%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[10pt,journal,compsoc]{../sty/IEEEtran}

@@

\definecolor{lightred}{rgb}{1.0,0.6,0.6}
\definecolor{lightgreen}{rgb}{0.6,1.0,0.6}

\DeclareRobustCommand{\add}[1]{{\sethlcolor{lightgreen}\hl{#1}}}
\DeclareRobustCommand{\remove}[1]{{\sethlcolor{lightred}\hl{#1}}}

\DeclareRobustCommand{\addCite}[1]{{\sethlcolor{lightgreen}\hl{\mbox{#1}}}}
\DeclareRobustCommand{\removeCite}[1]{{\sethlcolor{lightred}\hl{\mbox{#1}}}}

@@

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
    \begin{abstract}
        Numerical stability is a crucial requirement of reliable scientific
        computing. However, despite the pervasiveness of Python in data science,
        analyzing large Python programs remains challenging due to the lack of
        scalable numerical analysis tools available for this language. To fill
        this gap, we developed \pytracer, a profiler to quantify numerical
        instability in Python applications. \pytracer transparently instruments
        Python code to produce numerical traces and visualize them interactively
        in a Plotly dashboard. We designed \pytracer to be agnostic to numerical
        noise model , allowing for tool evaluation through Monte-Carlo
        Arithmetic, random rounding, random data perturbation, or structured
        noise for a particular application. We illustrate \pytracer's
        capabilities by testing the numerical stability of key functions in both
        SciPy and Scikit-learn, two dominant Python libraries for mathematical
        modeling. Through these evaluations, we demonstrate \pytracer as a
        scalable, automatic, and generic framework for numerical profiling in
        Python.
    \end{abstract}
}
% Note that keywords are not normally used for peerreview papers.
%% \begin{IEEEkeywords}
%% Computer Society, IEEE, IEEEtran, journal, \LaTeX, paper, template.
%% \end{IEEEkeywords}}

@@

We present PyTracer, a numerical stability profiling and visualization tool for
Python programs. \pytracer adopts a dynamic approach that evaluates numerical
stability using program execution traces, and is applicable to large programs
without manual intervention. The transparent nature of \pytracer is particularly
valuable in contrast to theoretical approaches based on condition numbers or
backward error analysis~\addCite{\cite{higham2002accuracy}}, which require detailed modeling of the program and its
numerical implementation. Similarly, static code analysis techniques~\addCite{\cite{goubault2006static}} hardly
scale to large code bases, particularly in dynamically-typed languages such as
Python.

\pytracer estimates the variability of floating-point variables by combining
program traces obtained with different numerical perturbations, which requires
(\add{a}\remove{1}) generating numerically-perturbed executions, (\add{b}\remove{2}) tracing floating-point
computations along program executions, and (\add{c}\remove{3}) visualizing stability
evaluations. 
\add{PyTracer uses the ``fuzzy" Python environment instrumented with Monte-Carlo
arithmetic to address the requirement (a). \mbox{\pytracer} addresses the
remaining requirements with (b) a dynamic instrumentation of Python functions,
modules, and classes through meta-programming, (c) an interactive Plotly
dashboard to visualize stability measures throughout program executions.
\mbox{\pytracer} is available at
\mbox{\href{https://github.com/big-data-lab-team/pytracer.git}{github.com/big-data-lab-team/pytracer.git}}
under MIT License.}
\remove{\mbox{\pytracer} addresses these requirements with (1) a ``fuzzy" Python
    interpreter instrumented with Monte-Carlo arithmetic, (2) a dynamic
    instrumentation of Python functions, modules, and classes through
    meta-programming, (3) an interactive Plotly dashboard to visualize stability
    measures throughout program executions. }

@@

\pytracer dynamically instruments Python modules without requiring source code
modifications in the application. To do so, it creates a new, instrumented
version of each module that preserves its original attributes so that the
application can transparently use it.
% \gkmod{}{Note: this form of monkey patching is basically global and
% transparent decorating, right?} \Yohan{Yes}
By default, \pytracer instruments all the attributes in a given module except
the special ones of the form \texttt{\_\_*\_\_}, which is common for reserved
Python methods. However, some attributes are not writable. In this case,
\pytracer \remove{preservers} \add{preserves} the original attributes and warns the user. The user can
also restrict the list of traced attributes through an inclusion/exclusion
mechanism. The main advantage of this instrumentation approach is the lack of
required modification in the application source code, in contrast with
\remove{decorator-based approaches} \add{a decorator-based approache~\mbox{\cite{hunt2019decorators}}}. 
In addition, this technique is scalable and
applicable in read-only environments such as Singularity containers or HPC
clusters.

@@

\remove{Listings~\mbox{\ref{fig:wrapper_creation}} and~\mbox{\ref{fig:generic_wrapper}} show
    \mbox{\pytracer}'s function instrumentation mechanism.}
For each function to instrument, \pytracer's \texttt{loader} class creates a
wrapper function, dynamically compiles it with the \texttt{compile} builtin
function, and substitutes it for the original module function.  Although simple,
this technique does not apply to callable class instances, i.e., class instances
that have a \texttt{\_\_call\_\_} attribute and are not of type
\texttt{FunctionType}. Indeed, applying the wrapper technique would 1) modify
the type of the class instance to \texttt{FunctionType}, which could cause
syntax and semantic errors, and 2) mask class attributes, leading to
\texttt{AttributeError} exceptions. \pytracer overrides the
\texttt{\_\_call\_\_} attribute with the wrapper function when possible to
overcome this issue. When the \texttt{\_\_call\_\_} attribute is not writable,
Pytracer does not instrument the class and returns a warning.

% \begin{listing}
%     \centering
%     \begin{lstlisting}[language=Python,style=customPython]
% def get_wrapper_function(module, qualname, name, function):
%     """Returns the instrumented function as a string.
%     This string will be compiled with the compile builtin function

%     Parameters:
%         module: Name of the function module
%         qualname: Qualified name of the function
%         name: Name of the function
%         function: Python object function

%     Returns:
%         wrapper_code: The code source of the wrapper
%     """
%     info = (id(function), function_module, qualname)
%     wrapper_code = f"""
%     def {function_wrapper_name}(*args, **kwargs):
%         return generic_wrapper({info},*args,**kwargs)"""
%     return wrapper_code
% \end{lstlisting}
%     \caption{Function to create the instrumented version of a function.
%         \pytracer uses the identifier of the function instead of the 
%         actual function to handle aliases and avoid duplicate instrumentation.}
%     \label{fig:wrapper_creation}
% \end{listing}

% \begin{listing}
%     \centering
%     \begin{lstlisting}[language=Python,style=customPython,]
% def generic_wrapper(self, info, *args, **kwargs):
%     """Generic wrapper saving inputs and outputs of the wrapped function.
%     The id_dict dict keeps a mapping between the original function identifier
%     and itself. Arguments are binded to avoid mispositioning.
%     Returns the actual output of the wrapped function.

%     Parameters:
%         info: Tuple with the function's id, function's module and function's name
%         *args: Positional calling arguments
%         **kwargs: Keyword calling arguments

%     Returns:
%         outputs: Outputs of the wrapped function
%     """
%     fid, module_name, function_name = info
%     function = original_function_cache.id_dict[fid]
%     bind = Binding(function, *args, **kwargs)
%     stack = self.backtrace()
%     time = elements()
%     self.inputs(time, module_name, function_name, function, bind.arguments, stack)
%     outputs = function(*bind.args, **bind.kwargs)
%     self.outputs(time, module_name, function_name, function, outputs, stack)
%     return outputs
% \end{lstlisting}
%     \caption{\pytracer's generic wrapper function that saves
%         the inputs and outputs of the wrapped function.}
%     \label{fig:generic_wrapper}
% \end{listing}

@@

\pytracer detects numerical instabilities by computing summary statistics across
multiple executions perturbed with numerical noise. While \pytracer's
instrumentation, trace aggregation, and visualization work with various types of
numerical noise, we experimented primarily with Monte-Carlo Arithmetic (MCA) as
it is an accurate model for floating-point errors~\addCite{\cite{verificarlo}}.

@@

Three  main  approaches  exist  to  detect  numerical  instabilities:
stochastic  arithmetic,  uncertainty  or  sensitivity analysis [23], and random
seed analysis [11].  \hl{PyTracer is agnostic to the noise model used and therefore
    works for all methods mentioned}.  For our use cases, we focus on stochastic
arithmetic since it does not make any assumption about the noise model, in
contrast to sensitivity analysis\add{.}

@@

Once traces are available, \pytracer sequentially parses them and computes the
mean, standard deviation, and the number of significant digits (as in
Equation~\ref{eq:sig-digits}) of all saved values. \pytracer saves these summary
statistics in an HDF5 file, a hierarchical format that facilitates visual
browsing since only required information are loaded when requested by the user.
This operation critically relies on the ordering of call input and output
objects in the traces. Therefore, it assumes that the application is
single-threaded and that the control flow does not depend on a random state. 
\add{Therefore, \mbox{\pytracer} cannot merge traces resulting from a different control flow path
    and will stop the parsing when encountering one.}

@@

We evaluated \pytracer on two widely-used Python libraries:
SciPy~\cite{virtanen2020scipy}, the main library for scientific computing in
Python, and  scikit-learn~\cite{pedregosa2011scikit}, a reference
machine-learning framework that implements a wide range of state-of-the-art
algorithms. For each library, we executed the software examples provided in
their respective GitHub repositories. We traced each example five times, having
activated MCA in the Python interpreter \add{as well as} BLAS, LAPACK, NumPy, SciPy, and
scikit-learn. \add{Only instrumented the Python interpreter would lead to
    only perturbe pure Python floating-point operations and thus to miss
    operations in compiled codes (i.e. Cython and shared libraries).}
\hl{We repeated the experiment with two MCA modes, Random Rounding
    (RR) and Full MCA, resulting in 2 experiments for each example}. We used the
default virtual precision in Verificarlo, namely 24 bits for single-precision
floating-point values and 53 bits for double-precision values, which simulates
machine error. We used Python-3.8.5, NumPy-1.19.1, SciPy-1.5.4, and
scikit-learn-0.23.2.

@@

\begin{listing}
    \begin{lstlisting}[language=C,style=customC,numbers=left, firstnumber=3163, mathescape=true]
/* NUMPY_API Arange, */
NPY_NO_EXPORT PyObject *
PyArray_Arange(double start, double stop, double step, int type_num) {
    npy_intp length; 
    PyArrayObject *range; $\lstsetnumber{\ldots}$
    ...$\lstresetnumber\setcounter{lstnumber}{3173}$
    double delta, tmp_len; $\lstsetnumber{\ldots}$
    ...$\lstresetnumber\setcounter{lstnumber}{3176}$
    delta = stop - start; 
    tmp_len = delta/step; $\lstsetnumber{\ldots}$
    ...$\lstresetnumber\setcounter{lstnumber}{3189}$
        length = _arange_safe_ceil_to_intp(tmp_len); $\lstsetnumber{\ldots}$
    ...$\lstresetnumber\setcounter{lstnumber}{3201}$
    range = (PyArrayObject *)PyArray_New(&PyArray_Type, 1, &length, type_num, NULL, NULL, 0, 0, NULL);
    \end{lstlisting}
    \caption{Source code of the \texttt{PyArray\_Range} Cython function called by NumPy function \texttt{linspace} to create an array of equally-spaced elements. The temporary array size assigned in line 3178 is stored as a floating-point value and is therefore perturbed in Full MCA mode, leading to differences amplified by the use of ceil rounding at line 3190 and resulting in different array sizes across MCA-perturbed executions. See also Table~\ref{tab:mca_result_linspace}.}
    \label{fig:pyarray_range_code}

@@

\add{The} top row in Figures~\ref{fig:fft1D_inputs} and~\ref{fig:fft1D_outputs} show the
input mean value while the bottom row shows the standard deviation. The points
with low magnitude in Figure~\ref{fig:fft1D_inputs} correspond to inputs near 0
when the input of $\sin$ or $\exp$ is close to $k\pi$, $k \in \mathbb{Z}$.
Indeed, since MCA introduces a slight noise, the result is not exactly 0. We can
see in Figure~\ref{fig:fft1D_inputs} that the maximal noise introduced by MCA on
the inputs is in the order of $10^{-14}$, two orders of magnitude higher than
the $ulp \simeq 10^-16$ for double precision. This slight noise introduced can
be interpreted as white noise on the input signal. We can see on the bottom row
of Figure~\ref{fig:fft1D_outputs} that the frequency noise is of the same order
of of magnitude as the input noise, which is expected. The two peaks at x=38 and
x=562 corresponds to the actual frequencies of the input signal.

@@

\paragraph{Unconstrained minimization of multivariate scalar functions}
Four Quasi-Newton-Raphson methods minimize function $f$ by using Newton's step:
$x_{k+1} = x_{k} - B^{-1}(x_k)\nabla f(x_k)$, where $\nabla$ and $B$ are the
gradient and the approximation of the Hessian matrix of $f$. The
\texttt{Broyden-Fletcher-Goldfarb-Shanno}~\cite{BFGS} (BFGS) and
\texttt{Newton-Conjugate-Gradient}~\cite{nocedal2006numerical} (NCG) are line
search methods that approximate $H$ by adding two symmetric rank-one matrices
and by using the Conjugate-Gradient method. The \texttt{Trust-Region Truncated
    Generalized Lanczos / Conjugate Gradient}~\cite{gould1999solving} (TR-Krylov),
\texttt{Trust-Region Newton-Conjugate-Gradient} (TR-NCG), and
\texttt{Trust-Region Nearly Exact}~\cite{nocedal2006numerical} (TR-Exact) are
trust-region methods approximating $H$ by solving the trust-region subproblem
restricted to a truncated Krylov subspace and by solving nonlinear equations for
each quadratic subproblem. In addition, the
\texttt{Nelder-Mead}~\cite{singer2009nelder} simplex-based method iteratively
transforms a simplex until its vertices are getting closer to a local minima of
the function. All methods are applied to the Rosenbrock function.

\paragraph{Constrained minimization of multivariate scalar functions}
Two optimizers solve an optimization problem by taking into account constraints.
The \texttt{Sequential Least SQuares Programming} uses the
SLSQP~\cite{kraft1988software} method to optimize the Rosenbrock function. The
\texttt{Least-squares minimization} uses the Trust Region Reflective
algorithm~\cite{li1993centering} to solve a fitting problem from an enzymatic
reaction.

The SLSQP and unconstrained minimization methods minimize the
\textbf{Rosenbrock} function of $N$ variables that reaches its minimum (0) when
$x_i=1,\; \forall i \leq N-1$:
$f(x) = \sum_{i=1}^{N-1} 100(x_{i+1}-x^2_i)^2 + (1-x_i)^2$.
Figure~\ref{fig:unconstrained_optimization} shows the number of significant bits
of the solution for the different methods used. The $(p)$ notation for the NCG,
Trust-Krylov, and Trust-NCG methods in the legend refers to the variant where
the Hessian matrix is replaced by a function computing the matrix-vector product
of the Hessian and an arbitrary vector, to save memory and time. All the methods
have a good precision ranging from 43 to 53 significant bits. The Nelder-Mead
method is the least precise one with $\simeq$ 43 significant bits on average,
while the Trust-Region-Newton-CG is the most precise one with $\simeq$ 53
significant bits, the highest precision achievable in double precision. The
remaining methods have a similar precision.

@@

\paragraph{Root finding}
% We tested three root finder algorithms with one for small problems using the hybrid Powell and Levenberg-Marquardt methods  (\texttt{root\_finding}) and two for large problems using a Krylov approximation for the inverse Jacobian with and without the help of a preconditioner (\texttt{root\_finding\_large\_preconditionned} and \texttt{root\_finding\_large}).

The \texttt{root\_*} examples use three algorithms to find the root of
non-linear equations. \texttt{root\_finding} uses the hybrid Powell method to
solve a single-variable transcendental equation and the Levenberg-Marquardt
method for a set of non-linear equations. \texttt{root\_finding\_large} and
\texttt{root\_finding\_large\_preconditioned} examples use the Krylov method to
approximate the inverse Jac\add{o}bian with and without the help of a preconditioner.
They solve an integrodifferential equation on a square $(x,y) \in [0,1] \times
    [0,1]$, $P(x,1)=1$, and $P=0$ elsewhere on the boundary. The function $P$ is
discretized with a Cartesian grid $P_{n,n}=P(nh,nh)$, with $n=75$, and its
derivatives are approximated by $\partial_x^2P(x,y) \simeq (P(x+h,y) - 2P(x,y) +
    P(x-h,y))/h^2$.

@@

This example compares a Bayesian Ridge Regression (BRR) to the Ordinary Least
Squares (OLS) estimator on a synthetic dataset and for one-dimensional
regression using polynomial feature expansion. Although the example does not
raise a runtime error, the regression coefficients computed from the fitting are
non-significant. \pytracer traces reveal that the SVD solver used is the root
cause for this error. Figure~\ref{fig:call_path_brr} shows the two call paths
for BRR and OLS. The LAPACK library has two main methods for computing the SVD:
\texttt{gesdd} and \texttt{gesvd}. The former uses a Divide \& Conquer approach,
while the latter uses a QR decomposition. While both methods are expected to
have the same accuracy~\cite{nakatsukasa2013stable},
Figure~\ref{fig:brr_svd_sig} shows that \texttt{gesdd} is totally imprecise,
with an average number of significant digits of 0 for BRR and even
\texttt{Not-a-Number} (\texttt{NaN}) values for OLS. On the same figure, we can
see that using the \texttt{gesvd} (\remove{dgelsd} \add{dgelss}) method instead of the \texttt{gesdd}
(\remove{dgelss} \add{dgelsd}) one, the SVD converges for RR and Full MCA, with a number of
significant bits of 44 on average. This observation supports the on-going
discussion among LAPACK developers on the instability of \texttt{gesdd} (see
\href{https://github.com/Reference-LAPACK/lapack/issues/316}{github.com/Reference-LAPACK/lapack/issues/316}).

@@

\bibliographystyle{IEEEtran}
\bibliography{paper-annotated}

>>>>

\usepackage{todonotes}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{color}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{xcolor,colortbl}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue!50!black,citecolor=blue!75!black,filecolor=black,urlcolor=blue!75!black}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage{makecell}
\usepackage{ulem} % provides sout
\usepackage[frozencache=true,cachedir=minted-cache]{minted} 
\usepackage[htt]{hyphenat}      
\usepackage{subcaption}
\usepackage{graphicx,import}
\usepackage{booktabs}
\usepackage{soul}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[10pt,journal,compsoc]{../sty/IEEEtran}

@@

\newcommand{\add}[1]{\colorbox{green!25!white}{#1}}
\newcommand{\remove}[1]{\colorbox{red!25!white}{\st{#1}}}

@@

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
    \begin{abstract}
        \add{Numerical stability is} a \remove{crucial requirement} of reliable scientific
        computing. However, despite the pervasiveness of Python in data science,
        analyzing large Python programs remains challenging due to the lack of
        scalable numerical analysis tools available for this language. To fill
        this gap, we developed \pytracer, a profiler to quantify numerical
        instability in Python applications. \pytracer transparently instruments
        Python code to produce numerical traces and visualize them interactively
        in a Plotly dashboard. We designed \pytracer to be agnostic to numerical
        noise model , allowing for tool evaluation through Monte-Carlo
        Arithmetic, random rounding, random data perturbation, or structured
        noise for a particular application. We illustrate \pytracer's
        capabilities by testing the numerical stability of key functions in both
        SciPy and Scikit-learn, two dominant Python libraries for mathematical
        modeling. Through these evaluations, we demonstrate \pytracer as a
        scalable, automatic, and generic framework for numerical profiling in
        Python.
    \end{abstract}
}
% Note that keywords are not normally used for peerreview papers.
%% \begin{IEEEkeywords}
%% Computer Society, IEEE, IEEEtran, journal, \LaTeX, paper, template.
%% \end{IEEEkeywords}}

@@

We present PyTracer, a numerical stability profiling and visualization tool for
Python programs. \pytracer adopts a dynamic approach that evaluates numerical
stability using program execution traces, and is applicable to large programs
without manual intervention. The transparent nature of \pytracer is particularly
valuable in contrast to theoretical approaches based on condition numbers or
backward error analysis, which require detailed modeling of the program and its
numerical implementation. Similarly, static code analysis techniques hardly
scale to large code bases, particularly in dynamically-typed languages such as
Python.

\pytracer estimates the variability of floating-point variables by combining
program traces obtained with different numerical perturbations, which requires
(1) generating numerically-perturbed executions, (2) tracing floating-point
computations along program executions, and (3) visualizing stability
evaluations. \pytracer addresses these requirements with (1) a ``fuzzy" Python
interpreter instrumented with Monte-Carlo arithmetic, (2) a dynamic
instrumentation of Python functions, modules, and classes through
meta-programming, (3) an interactive Plotly dashboard to visualize stability
measures throughout program executions.

@@

\pytracer dynamically instruments Python modules without requiring source code
modifications in the application. To do so, it creates a new, instrumented
version of each module that preserves its original attributes so that the
application can transparently use it.
% \gkmod{}{Note: this form of monkey patching is basically global and
% transparent decorating, right?} \Yohan{Yes}
By default, \pytracer instruments all the attributes in a given module except
the special ones of the form \texttt{\_\_*\_\_}, which is common for reserved
Python methods. However, some attributes are not writable. In this case,
\pytracer preservers the original attributes and warns the user. The user can
also restrict the list of traced attributes through an inclusion/exclusion
mechanism. The main advantage of this instrumentation approach is the lack of
required modification in the application source code, in contrast with
decorator-based approaches. In addition, this technique is scalable and
applicable in read-only environments such as Singularity containers or HPC
clusters.

@@

Listings~\ref{fig:wrapper_creation} and~\ref{fig:generic_wrapper} show
\pytracer's function instrumentation mechanism. For each function to instrument,
\pytracer's \texttt{loader} class creates a wrapper function, dynamically
compiles it with the \texttt{compile} builtin function, and substitutes it for
the original module function.  Although simple, this technique does not apply to
callable class instances, i.e., class instances that have a
\texttt{\_\_call\_\_} attribute and are not of type \texttt{FunctionType}.
Indeed, applying the wrapper technique would 1) modify the type of the class
instance to \texttt{FunctionType}, which could cause syntax and semantic errors,
and 2) mask class attributes, leading to \texttt{AttributeError} exceptions.
\pytracer overrides the \texttt{\_\_call\_\_} attribute with the wrapper
function when possible to overcome this issue. When the \texttt{\_\_call\_\_}
attribute is not writable,  Pytracer does not instrument the class and returns a
warning.

\begin{listing}
    \centering
    \begin{lstlisting}[language=Python,style=customPython]
def get_wrapper_function(module, qualname, name, function):
    """Returns the instrumented function as a string.
    This string will be compiled with the compile builtin function

    Parameters:
        module: Name of the function module
        qualname: Qualified name of the function
        name: Name of the function
        function: Python object function

    Returns:
        wrapper_code: The code source of the wrapper
    """
    info = (id(function), function_module, qualname)
    wrapper_code = f"""
    def {function_wrapper_name}(*args, **kwargs):
        return generic_wrapper({info},*args,**kwargs)"""
    return wrapper_code
\end{lstlisting}
    \caption{Function to create the instrumented version of a function.
        \pytracer uses the identifier of the function instead of the 
        actual function to handle aliases and avoid duplicate instrumentation.}
    \label{fig:wrapper_creation}
\end{listing}

\begin{listing}
    \centering
    \begin{lstlisting}[language=Python,style=customPython,]
def generic_wrapper(self, info, *args, **kwargs):
    """Generic wrapper saving inputs and outputs of the wrapped function.
    The id_dict dict keeps a mapping between the original function identifier
    and itself. Arguments are binded to avoid mispositioning.
    Returns the actual output of the wrapped function.

    Parameters:
        info: Tuple with the function's id, function's module and function's name
        *args: Positional calling arguments
        **kwargs: Keyword calling arguments

    Returns:
        outputs: Outputs of the wrapped function
    """
    fid, module_name, function_name = info
    function = original_function_cache.id_dict[fid]
    bind = Binding(function, *args, **kwargs)
    stack = self.backtrace()
    time = elements()
    self.inputs(time, module_name, function_name, function, bind.arguments, stack)
    outputs = function(*bind.args, **bind.kwargs)
    self.outputs(time, module_name, function_name, function, outputs, stack)
    return outputs
\end{lstlisting}
    \caption{\pytracer's generic wrapper function that saves
        the inputs and outputs of the wrapped function.}
    \label{fig:generic_wrapper}
\end{listing}

@@

\pytracer detects numerical instabilities by computing summary statistics across
multiple executions perturbed with numerical noise. While \pytracer's
instrumentation, trace aggregation, and visualization work with various types of
numerical noise, we experimented primarily with Monte-Carlo Arithmetic (MCA) as
it is an accurate model for floating-point errors.

@@

Three  main  approaches  exist  to  detect  numerical  instabilities:
stochastic  arithmetic,  uncertainty  or  sensitivity analysis [23], and random
seed analysis [11].  PyTracer is agnostic to the noise model used and therefore
works for all methods mentioned.  For our use cases, we focus on stochastic
arithmetic since it does not make any assumption about the noise model, in
contrast to sensitivity analysis

@@

Once traces are available, \pytracer sequentially parses them and computes the
mean, standard deviation, and the number of significant digits (as in
Equation~\ref{eq:sig-digits}) of all saved values. \pytracer saves these summary
statistics in an HDF5 file, a hierarchical format that facilitates visual
browsing since only required information are loaded when requested by the user.
This operation critically relies on the ordering of call input and output
objects in the traces. Therefore, it assumes that the application is
single-threaded and that the control flow does not depend on a random state.

@@

We evaluated \pytracer on two widely-used Python libraries:
SciPy~\cite{virtanen2020scipy}, the main library for scientific computing in
Python, and  scikit-learn~\cite{pedregosa2011scikit}, a reference
machine-learning framework that implements a wide range of state-of-the-art
algorithms. For each library, we executed the software examples provided in
their respective GitHub repositories. We traced each example five times, having
activated MCA in the Python interpreter, BLAS, LAPACK, NumPy, SciPy, and
scikit-learn. We repeated the experiment with two MCA modes, Random Rounding
(RR) and Full MCA, resulting in 2 experiments for each example. We used the
default virtual precision in Verificarlo, namely 24 bits for single-precision
floating-point values and 53 bits for double-precision values, which simulates
machine error. We used Python-3.8.5, NumPy-1.19.1, SciPy-1.5.4, and
scikit-learn-0.23.2.

@@

\begin{listing}
    \begin{lstlisting}[language=C,style=customC,numbers=right, firstnumber=3163, mathescape=true]
/* NUMPY_API Arange, */
NPY_NO_EXPORT PyObject *
PyArray_Arange(double start, double stop, double step, int type_num) {
    npy_intp length; 
    PyArrayObject *range; $\lstsetnumber{\ldots}$
    ...$\lstresetnumber\setcounter{lstnumber}{3173}$
    double delta, tmp_len; $\lstsetnumber{\ldots}$
    ...$\lstresetnumber\setcounter{lstnumber}{3176}$
    delta = stop - start; 
    tmp_len = delta/step; $\lstsetnumber{\ldots}$
    ...$\lstresetnumber\setcounter{lstnumber}{3189}$
        length = _arange_safe_ceil_to_intp(tmp_len); $\lstsetnumber{\ldots}$
    ...$\lstresetnumber\setcounter{lstnumber}{3201}$
    range = (PyArrayObject *)PyArray_New(&PyArray_Type, 1, &length, type_num, NULL, NULL, 0, 0, NULL);
    \end{lstlisting}
    \caption{Source code of the \texttt{PyArray\_Range} Cython function called by NumPy function \texttt{linspace} to create an array of equally-spaced elements. The temporary array size assigned in line 3178 is stored as a floating-point value and is therefore perturbed in Full MCA mode, leading to differences amplified by the use of ceil rounding at line 3190 and resulting in different array sizes across MCA-perturbed executions. See also Table~\ref{tab:mca_result_linspace}.}
    \label{fig:pyarray_range_code}

@@

Top row in Figures~\ref{fig:fft1D_inputs} and~\ref{fig:fft1D_outputs} show the
input mean value while the bottom row shows the standard deviation. The points
with low magnitude in Figure~\ref{fig:fft1D_inputs} correspond to inputs near 0
when the input of $\sin$ or $\exp$ is close to $k\pi$, $k \in \mathbb{Z}$.
Indeed, since MCA introduces a slight noise, the result is not exactly 0. We can
see in Figure~\ref{fig:fft1D_inputs} that the maximal noise introduced by MCA on
the inputs is in the order of $10^{-14}$, two orders of magnitude higher than
the $ulp \simeq 10^-16$ for double precision. This slight noise introduced can
be interpreted as white noise on the input signal. We can see on the bottom row
of Figure~\ref{fig:fft1D_outputs} that the frequency noise is of the same order
of of magnitude as the input noise, which is expected. The two peaks at x=38 and
x=562 corresponds to the actual frequencies of the input signal.

@@

\paragraph{Unconstrained minimization of multivariate scalar functions:}
Four Quasi-Newton-Raphson methods minimize function $f$ by using Newton's step:
$x_{k+1} = x_{k} - B^{-1}(x_k)\nabla f(x_k)$, where $\nabla$ and $B$ are the
gradient and the approximation of the Hessian matrix of $f$. The
\texttt{Broyden-Fletcher-Goldfarb-Shanno}~\cite{BFGS} (BFGS) and
\texttt{Newton-Conjugate-Gradient}~\cite{nocedal2006numerical} (NCG) are line
search methods that approximate $H$ by adding two symmetric rank-one matrices
and by using the Conjugate-Gradient method. The \texttt{Trust-Region Truncated
    Generalized Lanczos / Conjugate Gradient}~\cite{gould1999solving} (TR-Krylov),
\texttt{Trust-Region Newton-Conjugate-Gradient} (TR-NCG), and
\texttt{Trust-Region Nearly Exact}~\cite{nocedal2006numerical} (TR-Exact) are
trust-region methods approximating $H$ by solving the trust-region subproblem
restricted to a truncated Krylov subspace and by solving nonlinear equations for
each quadratic subproblem. In addition, the
\texttt{Nelder-Mead}~\cite{singer2009nelder} simplex-based method iteratively
transforms a simplex until its vertices are getting closer to a local minima of
the function. All methods are applied to the Rosenbrock function.

\paragraph{Constrained minimization of multivariate scalar functions:}
Two optimizers solve an optimization problem by taking into account constraints.
The \texttt{Sequential Least SQuares Programming} uses the
SLSQP~\cite{kraft1988software} method to optimize the Rosenbrock function. The
\texttt{Least-squares minimization} uses the Trust Region Reflective
algorithm~\cite{li1993centering} to solve a fitting problem from an enzymatic
reaction.

\paragraph{}
The SLSQP and unconstrained minimization methods minimize the
\textbf{Rosenbrock} function of $N$ variables that reaches its minimum (0) when
$x_i=1,\; \forall i \leq N-1$:
$f(x) = \sum_{i=1}^{N-1} 100(x_{i+1}-x^2_i)^2 + (1-x_i)^2$.
Figure~\ref{fig:unconstrained_optimization} shows the number of significant bits
of the solution for the different methods used. The $(p)$ notation for the NCG,
Trust-Krylov, and Trust-NCG methods in the legend refers to the variant where
the Hessian matrix is replaced by a function computing the matrix-vector product
of the Hessian and an arbitrary vector, to save memory and time. All the methods
have a good precision ranging from 43 to 53 significant bits. The Nelder-Mead
method is the least precise one with $\simeq$ 43 significant bits on average,
while the Trust-Region-Newton-CG is the most precise one with $\simeq$ 53
significant bits, the highest precision achievable in double precision. The
remaining methods have a similar precision.

@@

\paragraph{Root finding:}
% We tested three root finder algorithms with one for small problems using the hybrid Powell and Levenberg-Marquardt methods  (\texttt{root\_finding}) and two for large problems using a Krylov approximation for the inverse Jacobian with and without the help of a preconditioner (\texttt{root\_finding\_large\_preconditionned} and \texttt{root\_finding\_large}).

The \texttt{root\_*} examples use three algorithms to find the root of
non-linear equations. \texttt{root\_finding} uses the hybrid Powell method to
solve a single-variable transcendental equation and the Levenberg-Marquardt
method for a set of non-linear equations. \texttt{root\_finding\_large} and
\texttt{root\_finding\_large\_preconditioned} examples use the Krylov method to
approximate the inverse Jacbian with and without the help of a preconditioner.
They solve an integrodifferential equation on a square $(x,y) \in [0,1] \times
    [0,1]$, $P(x,1)=1$, and $P=0$ elsewhere on the boundary. The function $P$ is
discretized with a Cartesian grid $P_{n,n}=P(nh,nh)$, with $n=75$, and its
derivatives are approximated by $\partial_x^2P(x,y) \simeq (P(x+h,y) - 2P(x,y) +
    P(x-h,y))/h^2$.

@@

This example compares a Bayesian Ridge Regression (BRR) to the Ordinary Least
Squares (OLS) estimator on a synthetic dataset and for one-dimensional
regression using polynomial feature expansion. Although the example does not
raise a runtime error, the regression coefficients computed from the fitting are
non-significant. \pytracer traces reveal that the SVD solver used is the root
cause for this error. Figure~\ref{fig:call_path_brr} shows the two call paths
for BRR and OLS. The LAPACK library has two main methods for computing the SVD:
\texttt{gesdd} and \texttt{gesvd}. The former uses a Divide \& Conquer approach,
while the latter uses a QR decomposition. While both methods are expected to
have the same accuracy~\cite{nakatsukasa2013stable},
Figure~\ref{fig:brr_svd_sig} shows that \texttt{gesdd} is totally imprecise,
with an average number of significant digits of 0 for BRR and even
\texttt{Not-a-Number} (\texttt{NaN}) values for OLS. On the same figure, we can
see that using the \texttt{gesvd} (dgelsd) method instead of the \texttt{gesdd}
(dgelss) one, the SVD converges for RR and Full MCA, with a number of
significant bits of 44 on average. This observation supports the on-going
discussion among LAPACK developers on the instability of \texttt{gesdd} (see
\href{https://github.com/Reference-LAPACK/lapack/issues/316}{github.com/Reference-LAPACK/lapack/issues/316}).

@@

\bibliographystyle{IEEEtran}
\bibliography{paper}
